Biotic invasions reduce the manageability of mutualistic networks

### 2016-07-09 paper/manuscript.Rmd
add supplemetary info

---
title: "Biotic invasions reduce the manageability of mutualistic networks"
csl: ecology-letters.csl
output: html_document
bibliography: references.bib
---
```{r libraries, echo = FALSE, warning = F, message = F}
library(magrittr)
library(knitr)
library(ggplot2)
library(glmulti)
library(Hmisc)
library(lme4)
opts_knit$set (root.dir=normalizePath('../'))
```

```{r auxiliary functions, echo = FALSE, warning = F, message = F}
"./code/V2.0/functions" %>% 
	list.files(full.names = T) %>%
	plyr::l_ply(source)
```

@@

```{r read data and results, echo = FALSE, warning = F, message = F}
net <- "./data/V2.0/networks" %>%
	read_networks()
meta <- readr::read_csv("data/ntw_info.csv") %>% dplyr::tbl_df() %>%
	dplyr::rename(count = method)
n_matched <- readRDS("./data/V2.0/n_matched.rds")
a_m <- readRDS("./data/V2.0/n_matched_all_types.rds") 
r_w <- readRDS("./data/V2.0/random_weights.rds") %>% dplyr::tbl_df() %>%
  dplyr::rename_("net_name" = ".id")
r_d <- readRDS("./data/V2.0/random_directions.rds") %>% dplyr::tbl_df() %>%
  dplyr::rename_("net_name" = ".id")
s_p <- readRDS("./data/V2.0/species_properties.rds") %>% dplyr::tbl_df()
# mod <- readRDS("./data/V2.0/species_models.rds")
driver <- c("./data/V2.0/matching_frequency") %>% 
  list.files(full.names = T) %>%
  lapply(list.files, full.names = T) %>% 
  unlist() %>% 
  lapply(read_allRDS)
prop <- expand.grid(c("asymmetry", "max_dep"), 
                    c("scaled_FALSE", "scaled_TRUE")) %>% plyr::alply(1)
driver <- mapply(function(x, y){
  attr(x, "type") <- as.character(y[,1])
  attr(x, "scaled") <- stringr::str_split(as.character(y[,2]), "_") %>% unlist() %>% extract(2) %>% as.logical()
  return(x)
}, driver, prop)
driver_bi <- "./data/V2.0/matching_frequency_bi" %>% 
  read_allRDS()
mod <- readRDS("./data/V2.0/detailed_species_models.rds")
mo <- readRDS("./data/V2.0/detailed_species_models_data.rds")
```

@@

Center of Integrative Ecology, School of Biological Sciences, University of Canterbury, 8041 Christchurch, New Zealand

@@

In the last section, we outlined the theoretical and computational underpinnings of our approach. 
Now we illustrate its empirical application using ten pairs of plant-pollinator communities for which we explored their manageability and the relative importance of their constituent species. 
The networks were constructed from previously published visitation data collected from pollination communities in Bristol, UK [@Lopezaraiza-Mikel2007] and in Cap de Creus National Park, Spain [@Bartomeus2008]. 
Notably, each network pair was composed of a community invaded by a plant and a community ''free'' of the invasive species (Table S1). 
In the four British uninvaded communities, the non-invaded plots were obtained by experimentally removing all the flowers of the invasive species *Impatients grandulifera*. 
In contrast, the Spanish uninvaded communities were obtained from plots that had not yet been colonised by either of the invasive species *Carpobrotus affine acinaciformis* or *Opuntia stricta*. 
The structure of all networks was determined using visitation frequency which has been shown to be an appropriate surrogate for interspecific effects in pollination networks [@Vazquez2005; @Bascompte2006]. 
To examine whether this decision would influence our results, we also evaluated the effect of using pollinator efficiency or importance as alternative measures of species interactions in a different dataset [@NeEman2010; @Ballantyne2015], and we found quantitatively similar results for all of these options (see Supplementary Information).

@@

Mutual dependencies have previously been shown to be highly asymmetric in natural communities [@Bascompte2006]; in other words, if a plant species is largely dependent on a pollinator species, then that pollinator tends to depend rather weakly on the plant. 
This suggests that it may be possible to relax the assumption of bidirectional dependences and simplify the network further such that interacting species are only connected by one directed link when mutual dependencies are asymmetric (Figure 2c). 
The extent to which a species $i$ affects species $j$ relative to the extent to which species $j$ affects species $i$ can be summarised by the interaction asymmetry given by $$a(i,j) = \frac{\left | d_{ij}-d_{ji} \right |}{max\left ( d_{ij}, d_{ji} \right )}.$$
Employing an unidrectional interaction asymmetry, as a surrogate of bidirectional mutual dependences, confers an additional advantage for our purposes since it dramatically reduces the number of links in the network. 
This translates into a substantial reduction of the time and memory required for the computation of of all possible maximum matchings in a network, particularly for large networks (see ''relative importance of species'' subsection below).

@@

Our second key question is related to how species differ in their importance for driving the population dynamics of the community. 
To quantify this importance in each network, we computed all possible matchings that have the same size as the networks' maximum matching. 
We then calculated the frequency with which each species is deemed to be a superior node in the set of matchings that had a total weight greater or equal to 0.5 times the weight of the maximum matching. 
We selected this threshold as it provided a high agreement between networks quantified by visitation and pollination efficiency as well as between our weighting/directionality assumptions (mutual dependences and interaction asymmetries); however, the choice of this threshold had a negligible impact on the results (Supplementary Information).

@@

```{r species, include = FALSE, warning = FALSE, message = FALSE}
data <- mo
net <- "./data/V2.0/networks" %>%
	read_networks()

>>>>

---
title: "Biotic invasions reduce the manageability of mutualistic networks"
csl: ecology-letters.csl
output: html_document
bibliography: references.bib
---

```{r libraries, echo = FALSE, warning = F, message = F}
library(magrittr)
library(knitr)
library(ggplot2)
library(glmulti)
library(Hmisc)
library(lme4)
opts_knit$set (root.dir=normalizePath('../'))
```

```{r auxiliary functions, echo = FALSE, warning = F, message = F}
"./code/functions" %>% 
	list.files(full.names = T) %>%
	plyr::l_ply(source)
```

@@

```{r read data and results, echo = FALSE, warning = F, message = F}
net <- "./data/processed/networks" %>%
	read_networks()
meta <- readr::read_csv("data/ntw_info.csv") %>% dplyr::tbl_df() %>%
	dplyr::rename(count = method)
n_matched <- readRDS("./data/processed/n_matched.rds")
a_m <- readRDS("./data/processed/n_matched_all_types.rds") 
r_w <- readRDS("./data/processed/random_weights.rds") %>% dplyr::tbl_df() %>%
  dplyr::rename_("net_name" = ".id")
r_d <- readRDS("./data/processed/random_directions.rds") %>% dplyr::tbl_df() %>%
  dplyr::rename_("net_name" = ".id")
s_p <- readRDS("./data/processed/species_properties.rds") %>% dplyr::tbl_df()
# mod <- readRDS("./data/V2.0/species_models.rds")
driver <- c("./data/processed/matching_frequency") %>% 
  list.files(full.names = T) %>%
  lapply(list.files, full.names = T) %>% 
  unlist() %>% 
  lapply(read_allRDS)
prop <- expand.grid(c("asymmetry", "max_dep"), 
                    c("scaled_FALSE", "scaled_TRUE")) %>% plyr::alply(1)
driver <- mapply(function(x, y){
  attr(x, "type") <- as.character(y[,1])
  attr(x, "scaled") <- stringr::str_split(as.character(y[,2]), "_") %>% unlist() %>% extract(2) %>% as.logical()
  return(x)
}, driver, prop)
driver_bi <- "./data/processed/matching_frequency_bi" %>% 
  read_allRDS()
mod <- readRDS("./data/processed/detailed_species_models.rds")
mo <- readRDS("./data/processed/detailed_species_models_data.rds")
```

@@

Centre for Integrative Ecology, School of Biological Sciences, University of Canterbury, 8041 Christchurch, New Zealand

@@

In the last section, we outlined the theoretical and computational underpinnings of our approach. 
Now we illustrate its empirical application using ten pairs of plant-pollinator communities for which we explored their manageability and the relative importance of their constituent species. 
The networks were constructed from previously published visitation data collected from pollination communities in Bristol, UK [@Lopezaraiza-Mikel2007] and in Cap de Creus National Park, Spain [@Bartomeus2008]. 
Notably, each network pair was composed of a community invaded by a plant and a community ''free'' of the invasive species (Table S1). 
In the four British uninvaded communities, the non-invaded plots were obtained by experimentally removing all the flowers of the invasive species *Impatients grandulifera*. 
In contrast, the Spanish uninvaded communities were obtained from plots that had not yet been colonised by either of the invasive species *Carpobrotus affine acinaciformis* or *Opuntia stricta*. 
The structure of all networks was determined using visitation frequency which has been shown to be an appropriate surrogate for interspecific effects in pollination networks [@Vazquez2005; @Bascompte2006]. 
To examine whether this decision would influence our results, we also evaluated the effect of using pollinator efficiency or importance as alternative measures of species interactions in a different dataset [@NeEman2010; @Ballantyne2015], and we found quantitatively similar results for all of these options (Supplementary Information S1).

@@

Mutual dependencies have previously been shown to be highly asymmetric in natural communities [@Bascompte2006]; in other words, if a plant species is largely dependent on a pollinator species, then that pollinator tends to depend rather weakly on the plant. 
This suggests that it may be possible to relax the assumption of bidirectional dependences and simplify the network further such that interacting species are only connected by one directed link when mutual dependencies are asymmetric (Figure 2c). 
The extent to which a species $i$ affects species $j$ relative to the extent to which species $j$ affects species $i$ can be summarised by the interaction asymmetry given by $$a(i,j) = \frac{\left | d_{ij}-d_{ji} \right |}{max\left ( d_{ij}, d_{ji} \right )}.$$
Employing an unidrectional interaction asymmetry, as a surrogate of bidirectional mutual dependences, does not alter the relative manageabilities of different networks (Supplementary Information S2). 
Moreover, it confers an additional advantage for our purposes since it dramatically reduces the number of links in the network. 
This translates into a substantial reduction of the time and memory required for the computation of of all possible maximum matchings in a network, particularly for large networks (see ''relative importance of species'' subsection below).

@@

Our second key question is related to how species differ in their importance for driving the population dynamics of the community. 
To quantify this importance in each network, we computed all possible matchings that have the same size as the networks' maximum matching. 
We then calculated the frequency with which each species is deemed to be a superior node in the set of matchings that had a total weight greater or equal to 0.5 times the weight of the maximum matching. 
We selected this threshold as it provided a high agreement between networks quantified by visitation and pollination efficiency as well as between our weighting/directionality assumptions (mutual dependences and interaction asymmetries); however, the choice of this threshold had a negligible impact on the results (Supplementary Information S3).

@@

```{r species, include = FALSE, warning = FALSE, message = FALSE}
data <- mo
net <- "./data/processed/networks" %>%
	read_networks()
